\documentclass[11pt, a4paper]{article}
%\usepackage{geometry}
\usepackage[inner=1.5cm,outer=1.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\pagestyle{empty}
\usepackage{graphicx}

\usepackage[usenames,dvipsnames]{color}
\definecolor{darkblue}{rgb}{0,0,.6}
\definecolor{darkred}{rgb}{.7,0,0}
\definecolor{darkgreen}{rgb}{0,.6,0}
\definecolor{red}{rgb}{.98,0,0}
\usepackage[colorlinks,pagebackref,pdfusetitle,urlcolor=darkblue,citecolor=darkblue,linkcolor=darkred,bookmarksnumbered,plainpages=false]{hyperref}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}



\thispagestyle{plain}

%%%%%%%%%%%% LISTING %%%
\usepackage{listings}
\usepackage{caption}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
\usepackage{verbatim} % used to display code
\usepackage{fancyvrb}
\usepackage{acronym}
\usepackage{amsthm}
\VerbatimFootnotes % Required, otherwise verbatim does not work in footnotes!



\definecolor{OliveGreen}{cmyk}{0.64,0,0.95,0.40}
\definecolor{CadetBlue}{cmyk}{0.62,0.57,0.23,0}
\definecolor{lightlightgray}{gray}{0.93}



\lstset{
%language=bash,                          % Code langugage
basicstyle=\ttfamily,                   % Code font, Examples: \footnotesize, \ttfamily
keywordstyle=\color{OliveGreen},        % Keywords font ('*' = uppercase)
commentstyle=\color{gray},              % Comments font
numbers=left,                           % Line nums position
numberstyle=\tiny,                      % Line-numbers fonts
stepnumber=1,                           % Step between two line-numbers
numbersep=5pt,                          % How far are line-numbers from code
backgroundcolor=\color{lightlightgray}, % Choose background color
frame=none,                             % A frame around the code
tabsize=2,                              % Default tab size
captionpos=t,                           % Caption-position = bottom
breaklines=true,                        % Automatic line breaking?
breakatwhitespace=false,                % Automatic breaks only at whitespace?
showspaces=false,                       % Dont make spaces visible
showtabs=false,                         % Dont make tabls visible
columns=flexible,                       % Column format
morekeywords={__global__, __device__},  % CUDA specific keywords
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{center}
  {\Large \textsc{Problem Set 1}}

  MGMT 737
\end{center}
\begin{center}
  Spring 2024
\end{center}
%\date{September 26, 2014}

\begin{enumerate}
\item Randomization. This analysis will use the Dehijia
  and Wahba sample from the Lalonde dataset of the NSW experiment. The dataset is
  \texttt{lalonde\_nsw.csv}. The outcome variable is \texttt{re78}
  (real earnings in 1978). The treatment indicator is
  \texttt{treat}. The remaining variables are potential covariates. Assume for the purposes of this problem set that $\texttt{treat}$ is completely randomly assigned. 
  \begin{enumerate}
  \item Calculate the average treatment effect of the policy
    $E(\tau_{i})$ using a simple difference in means.
  \item Calculate the average treatment effect on the treated of the policy
    $E(\tau_{i}| \mathrm{treat} = 1)$. How does it compare to part a? 
  \item Test the null of $\tau_{i} = 0$ for all $i$ using a
    randomization test. \\
    \emph{N.B.} Hold fixed the number of treated and control
    (e.g. assume the treatment count would be held fixed) and permute
    the labels randomly 1000 times -- you do not need to fully do
    every permutation (there would be too many). Report the quantile
    that your estimate from the previous question falls.
  \item Run a regression using robust standard errors (you may use
    canned software) of the outcome on the treatment dummy, and
    compare the p-values from this test to the previous answer.
  \end{enumerate}  
\end{enumerate}

\end{document}