---
output:
    rmdformats::robobook:
        self_contained: true
        thumbnails: true
        lightbox: true
        gallery: false
        highlight: tango
title: "Problem Set 2 Solutions"
author: "Paul Goldsmith-Pinkham"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
We first load in the Dehijia and Wahba sample from the Lalonde dataset of the NSW experiment. The dataset is ``lalonde_nsw.csv``. We will only take the treated units from this sample. We will then also load the ``lalonde_psid.csv`` as the control sample.  The outcome variable is ``re78`` (real earnings in 1978). The remaining variables are potential covariates we will use for our propensity score estimation.

First we load some convenience packages and set seed:
```{r, message = FALSE}
library(tidyverse)
library(data.table)
library(fixest)
set.seed(1234)
```
## Part a
Using ``age, education, hispanic, black, married, nodegree, RE74`` and ``RE75``, construct a propensity score using the treated group in ``lalonde_nsw.csv`` and the control sample of ``lalonde_psid.csv``.  Use a logit regression model to do so (you may use a canned routine to run the regression). Report the average p-score for the treated and control samples, and plot the propensity score densities for the treatment and control groups.

```{r, message = FALSE}
nsw_data <- read_csv("data/lalonde_nsw.csv")
psid_data <- read_csv("data/lalonde_psid.csv")

nsw_treat <- nsw_data[nsw_data$treat == 1, ]
psid_control <- psid_data[psid_data$treat == 0, ]

dw_data <- rbind(nsw_treat, psid_control)
```
```{r}
logit_model <- feglm(treat ~ age + education + hispanic + black + married + nodegree + re74 + re75, data = dw_data, family = "binomial")
pscore_logit <- predict(logit_model, dw_data)

dw_data %>%
    bind_cols(tibble(pscore_logit = pscore_logit)) %>%
    group_by(treat) %>%
    summarize(pscore_logit = mean(pscore_logit))
```

```{r}
# Density Plot
ggplot(dw_data %>% bind_cols(tibble(pscore_logit = pscore_logit))) +
    geom_density(aes(x = pscore_logit, color = factor(treat), group = factor(treat))) +
    labs(x = "Propensity Score", y = "Density")
# Trimmed Density Plot
ggplot(dw_data %>% bind_cols(tibble(pscore_logit = pscore_logit)) %>% filter(pscore_logit > 0.1 & pscore_logit < 0.9)) +
    geom_density(aes(x = pscore_logit, color = factor(treat), group = factor(treat))) +
    labs(x = "Propensity Score", y = "Density")
```


## Part b
Using your p-score estimates, calculate the IPW and SIPW
    estimate for control and treated mean of the outcome, and the
    average treatment effect. Contrast these estimates to the control
    mean of the outcome from the NSW sample, and the treatment effect
    from last week's problem set.

```{r}
ipw_estimate <- function(ps_hat, y, treat) {
    control_mean <- mean(y * (1 - treat) / (1 - ps_hat))
    treatment_mean <- mean(y * treat / (ps_hat))
    ate <- treatment_mean - control_mean
    return(list(ate = ate, control_mean = control_mean, treatment_mean = treatment_mean))
}
sipw_estimate <- function(ps_hat, y, treat) {
    control_mean <- (mean(y * (1 - treat) / (1 - ps_hat)) / mean((1 - treat) / (1 - ps_hat)))
    treatment_mean <- (mean(y * treat / (ps_hat)) / mean(treat / ps_hat))
    ate <- treatment_mean - control_mean
    return(list(ate = ate, control_mean = control_mean, treatment_mean = treatment_mean))
}
ipw_estimate(pscore_logit, dw_data$re78, dw_data$treat)
sipw_estimate(pscore_logit, dw_data$re78, dw_data$treat)
```

## Part c
Compare the ATE in the previous question to the treatment
    effect estimated using a linear regression using the PSID and NSW
    treatment sample, with `age`, `education`,
    `hispanic`, `black`, `married`,
    `nodegree`, `RE74` and `RE75` as controls.

```{r}    
linear_model_outcome <- feols(re78 ~ treat + age + education + hispanic + black + married + nodegree + re74 + re75, data = dw_data)
summary(linear_model_outcome)
```

## Part d
Now revisit your estimates from part a and b, and following
    Crump et al. (2009), discard all units with estimated propensities
    outside the range of $[0.1, 0.9]$. Reestimate the IPW and SIPW
    estimator of the ATE from part b using this trimmed sample.

```{r}
subset <- pscore_logit <= 0.9 & pscore_logit >= 0.1
ipw_estimate(
    pscore_logit[subset],
    dw_data$re78[subset],
    dw_data$treat[subset]
)
sipw_estimate(
    pscore_logit[subset],
    dw_data$re78[subset],
    dw_data$treat[subset]
)
```

## part e
Finally, calculate the IPW and SIPW estimates for the ATE
    using this trimmed sample for Black and non-Black
    individuals. Compare this estimate to the ATE for Black and
    non-Black individuals using the full randomized sample.

Black Estimates:
```{r}
subset <- pscore_logit <= 0.9 & pscore_logit >= 0.1 & dw_data$black == 1
ipw_estimate(
    pscore_logit[subset],
    dw_data$re78[subset],
    dw_data$treat[subset]
)
sipw_estimate(
    pscore_logit[subset],
    dw_data$re78[subset],
    dw_data$treat[subset]
)
```

White Estimates:
```{r}
subset <- pscore_logit <= 0.9 & pscore_logit >= 0.1 & dw_data$black == 0
ipw_estimate(
    pscore_logit[subset],
    dw_data$re78[subset],
    dw_data$treat[subset]
)
sipw_estimate(
    pscore_logit[subset],
    dw_data$re78[subset],
    dw_data$treat[subset]
)
```

Compare to full randomized sample:
```{r}
estimate_mean_diff <- function(treat, y) {
    return(mean(y[treat == 1]) - mean(y[treat == 0]))
}

estimate_mean_diff(nsw_data$treat[nsw_data$black == 1], nsw_data$re78[nsw_data$black == 1])
estimate_mean_diff(nsw_data$treat[nsw_data$black == 0], nsw_data$re78[nsw_data$black == 0])
```


## Part f
We now consider what our results look like if we just consider difference-in-differences. Estimate a linear regression model (using canned software package like `lm` in R is fine -- be sure to cluster on the individual) with two specifications:
    \begin{enumerate}
      \item The difference in `RE78` less `RE75` on the left hand side against a treatment indicator. 
      \item `RE78` on the left hand side against a treatment indicator and `RE75`.
    \end{enumerate}
  Report the estimates and compare to your previous estimates. Can you think of other specifications that might work? 

```{r}
did_model_outcome1 <- feols(re78 - re75 ~ treat, data = dw_data)
did_model_outcome2 <- feols(re78 ~ treat + re75, data = dw_data)
summary(did_model_outcome1)
summary(did_model_outcome2)
```
