
---
output:
    rmdformats::robobook:
        self_contained: true
        thumbnails: true
        lightbox: true
        gallery: false
        highlight: tango
title: "Problem Set 5 Solutions"
author: "Paul Goldsmith-Pinkham"
---


```{r setup, include=FALSE}
library(fixest)
library(tidyverse)
```


# Problem 1 -- Poisson regs
```{r, include = FALSE}
detroit <- read_csv("data/Detroit.csv")
```


## Part a 
Run an OLS regression of `flows` on
    `distance\_Google\_miles`, and include `home\_ID` and
    `work\_ID` fixed effects (absorb them), and cluster on
    \texttt{home\_ID}. Report the coefficient and standard error on
    \texttt{distance\_Google\_miles}.
```{r}
feols(flows ~ distance_Google_miles | home_ID + work_ID, vcov=~home_ID, data = detroit)
```


## Part b
Run an OLS regression of \texttt{log(flows)} on
    \texttt{log(distance\_Google\_miles)} and include \texttt{home\_ID}
    and \texttt{work\_ID} fixed effects , omitting the cells with zero
    flows. Cluster on \texttt{home\_ID}. Report the coefficient and
    standard error on \texttt{log(distance\_Google\_miles)}.
```{r}
feols(log(flows) ~ log(distance_Google_miles), data = detroit  %>% filter(flows > 0))
```

## Part c

  \item Repeat part 1b, but instead of omitting the zero cells, run
    the OLS regression of \texttt{log(c + flows)} for c = 0.1, 1 and
    10. Compare how your coefficients change
```{r}
for (c in c(0.1, 1, 10)) {
  print(feols(log(flows + c) ~ log(distance_Google_miles), data = detroit))
}
```

We see very large changes in the coefficient, likely driven by the fact that there are a number of observations with zero flows. Hence, the choice of $c$ is extremely important, as pointed out in Chen and Roth. 


## Part d

  \item Finally, repeat part 1a using Poisson regression, and contrast
    the estimates to Part b and c.

```{r}
fepois(flows ~ distance_Google_miles | home_ID + work_ID, vcov=~home_ID, data = detroit)
```
We see similar estimated coefficients to part a, but recall that the interpretation of the Poisson regression is very different -- Poisson estimates $log(E(Y|X))$ rather than $E(Y|X)$. Hence, we can interpret the effect as `r round((exp(-0.8854) - 1)*100)` percent change in the expected value of flows for a one unit change in distance.

To have the same interpretation in part a, we would run
```{r}
feols(log(flows) ~ distance_Google_miles | home_ID + work_ID, vcov=~home_ID, data = detroit)
```
and get `r round((exp(-0.039045)-1)*100)` percent change in the expected value of flows for a one unit change in distance.

# Problem 2 -- Duration model

```{r}

acs_data <- read_csv("data/acs_duration.csv")

```
# part a
Using the ACS data, write down how to estimate the unconditional probability that a household stays in a home for T or more years, using the available data. Estimate this for $T=7$ and report the value.

We can observe the empirical CDF of the duration variable. 
```{r}
acs_data %>% 
ggplot(aes(x = moving_approx)) + 
  stat_ecdf() + 
  labs(title = "CDF of 'How long have you lived here?'")
```

The share of households that have lived in their home for 7 years or more is `r mean(acs_data$moving_approx >= 7)`.

However, this will give us a biased estimate of the share of households that live in their home for 7 years or more, as we are not accounting for censoring. 

# Part b
Calculate the hazard rate for each observed value of
      `moving\_approx`. Report this value for $T=7$.

# Part c
    \item Recalculate these hazard values for $T=7$ for homeowners and
      renters. Contrast the difference in hazard rates over time.